---
title: "decrypt"
author: "Ashwath Raj"
date: "May 10, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(coda)
library(stringr)
```

##Report:
I set up an energy function based on the pairwise frequencies. I use a for loop to show all the energies given substitution shifts from 0-25. I saw all the output and I don't see an immediate "solve" or a point where manual tweaks made sense.

I haven't implemented mcmc() yet, but I'm curious how it would change the result because it looks like possible states are the possible shifts I already tested.


```{r}
# read coded message
input <- readLines("CodedMessage_Short.txt", n=1000)
input2 <- readLines("CodedMessage_Med.txt", n=1000)
# read frequency table
p <-read.table("LetterPairFreqFrom7Novels.txt")
input <- tolower(input)
s <- strsplit(input, split = " ")
s <- do.call(paste0, as.data.frame(s))
s2 <- strsplit(input2, split = " ")
s2 <- do.call(paste0, as.data.frame(s2))

# p[,1]
# 
# for (word in s){
#     for(char in 1:str_length(word)){
#       print(char)
#     }
# }
      
#At a given substitution point, evaluate the energy
energy <- function(strings, pairWeight, shift){
  freqs = 0
  alphabet <- letters#[1:26]
  wls <- str_length(strings) #word lengths
  
  for (word in strings){
    for(char in 1:(str_length(word)-1)){
        # if (char < str_length(word)){ #Can't index past the last letter
         letter1 = (match(substr(word,char,char),alphabet) + shift) %% 26
         letter2 = (match(substr(word,char+1,char+1), alphabet) + shift) %% 26
         
         if(!is.na(letter1) & !is.na(letter2)){
           if(letter1 != 0 & letter2 != 0){
              freqs = freqs + (pairWeight[letter1,letter2])
              # print(freqs)
         }
      }
                # for (line in p){ #Every line of frequencies
        #   for (letter in 1:26){ #Every letter frequency
        #     
        #   }
        # }
    }
    # browser()
  }
  return(freqs)

}

#At a given substitution point, evaluate the energy
convert <- function(strings, shift){
  new_msg = ""
  for (word in strings){
    new_word = ""
      for(char in 1:(str_length(word))){
        # chars = paste(char,())
        letter_num = (match(substr(word, char, char),letters) + shift)
        if(!is.na(letter_num)){
          if(letter_num != 0){
          if (letter_num > 26){letter_num = letter_num %% 26}
          letter = letters[letter_num]
          new_word = paste(new_word, letter, sep = "")
          }
        }
      }
    new_msg = paste(new_msg, new_word, sep = " ")
  }
  return(new_msg)
}


#Finding the best energy for a shift for s
max = 0
best_shift = 0
for (shift_x in 0:(length(letters)-1)){
  fx = energy(s,p,shift_x)
  print(shift_x)
  print(fx)
  # print(convert(s,shift_x))
  if(fx > max){
    max = fx
    best_shift = shift_x
   }
}
print(convert(s,shift = best_shift))


#Finding the best energy for a shift for s2
max = 0
best_shift = 0
for (shift_x in 0:(length(letters)-1)){
  fx = energy(s,p,shift_x)
  print(shift_x)
  print(fx)
  # print(convert(s,shift_x))
  if(fx > max){
    max = fx
    best_shift = shift_x
   }
}
print(convert(s2,shift = best_shift))
```

